Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Managed to put it back to 8 pages
Closes #95
Closes #98

\numberofauthors{5} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
% \author{Juan-J.~Merelo, Dept. of Computer Architecture and Technology
%   and CITIC, University of Granada, Granada, Spain \\
% Mario Garc\'ia-Valdez, Dept. of Graduate Studies, Instituto
% Tecnológico de Tijuana, Tijuana, M\'exico\\
% Pedro A. Castillo, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
% Pablo Garc\'ia-S\'anchez, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
% Paloma de las Cuevas, Dept. of Computer Architecture and Technology,
% University of Granada, Granada, Spain\\
% Nuria Rico, Dept. of Statistics and Operational Research, University of Granada, Granada, Spain%

@@

Some time ago, our group faced the problem of diminishing funds for
buying new hardware. This was aggravated by the increasing maintenance costs and
extended downtime resulting from the continuous failures of existing
clusters.
Considering this, we leveraged our experience in the design of web
applications since their inception, including JavaScript, and other
volunteer and
unconventional distributed evolutionary computing systems to 
design and release a new free framework that would allow anyone to
create a volunteer distributed evolutionary computation (EC) experiment using cloud resources as
servers, and browsers as clients.
% (Paloma) It seems that you've started the introduction twice. The first two paragraphs "introduce the introduction", and so does the third. It is specially noticeable when you find "evolutionary computation (EC)" twice.

@@

%Mario: Rewrote: 
%One of the aspects that must be taken into account in volunteer based
%frameworks is its {\em social} aspect in the design, with 
In general, several volunteers will be needed to
overcome the overhead involved in using a client-server architecture
and this number of volunteers is obviously not guaranteed. Therefore, the {\em social} aspect must be taken into account when designing 
volunteer based frameworks, with 
issues related to security, trust, and privacy among others. 
The computing system becomes then a {\em techno-social system}
\cite{vespignani2009predicting}, in the sense that its social aspects
cannot be separated from its technical specifications, with one
feeding into the other. One of these aspects is safety for the user running it, as well as for the scientist: the user must be reasonably sure that the
experiment is not misusing their resources and the scientist must have a
certain safety in the sense that their server is not going to be
attacked via cheating or denial of service attacks. We have decided to
approach this in a pragmatic way. Including many security features % (Paloma) I'm kind of lost here... Sentences seem to be unconnected and unrelated...
would decrease performance; on the other hand, we publish all data
resulting from the experiment once private information, such as IPs,
has been processed to an untraceable hash. The experiments themselves
are a proof that there are enough users trusting the experiment to
give it their CPU time and that there has been no attempt to cheat by
sending specially crafted chromosomes or to attack the server to make
it stop working because there was no incentive in doing so.
% (Paloma) Is this really true? Can we really state that there were no attacks because there was no incentive? Attackers don't always need an incentive to perform an attack, or the incentive could not be perceived by us. And why do the experiments prove this? Are there specific measures?
The
system, besides, is running in a dedicated Platform as a Service (PaaS)
container or Virtual Private Server so accessing the server would only
show the few data obtained from the experiment, adding security against attackers. In addition, making it
free software ensured this was known by anyone in the experiment, via
links in the web page that hosted it. % (Paloma) What was known? The fact that it was running in a PaaS? Why was that known by links?

In this paper, we will carry out three experiments, after establishing
the baseline of the performance that can be expected in a JavaScript
based system and its comparison with other languages in
\cite{2016arXiv160101607Manom}. First, several experiments are performed using a PaaS, and a simple Ajax client that does not
restart when a solution is found. Second, we change the client a bit so that it uses web workers, that is, it runs in the background process and then it restarts. Third, we use the same
% Pablo: it restarts when the solution is found (it has not been mentioned yet)
client, but with a harder problem that cannot be easily solved by a single
volunteer and thus needs the contribution of many volunteers to
proceed. 
%Mario: refrased: 
%With these three experiments we want mainly to try and model
With these three experiments our main objective is to try and model
the behavior of users and learn how different problems and
implementation choices have an influence on it. That is why in this
paper we will collect information on the number of users participating
in each experiment and also the time they spent doing so. In this
paper we will show how this participation follows certain patterns,
and that they depend on the implementation itself as well as the
problem and its difficulty; however, certain implementations allow to
extract more CPU cycles from volunteers.

@@

%
% \begin{figure}[!htb]
% \centering
% \includegraphics{rastrigin-fitness.png}
% \caption{Fitness of chromosomes sent to the server pool along time in
%   the $x$ axis.} 
% % Plotted with ../data/plot-rastrigin-fitness.R
% \label{fig:puts:rastrigin}
% \end{figure}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{rastrigin-IPs.png}
\includegraphics[width=0.49\linewidth]{rastrigin-IPs-hour.png}
\caption{Number of unique IPs participating in the Rastrigin
  experiment per minute (left) and per hour (right)} 
% Plotted with ../data/plot-rastrigin-IPs.R
\label{fig:ips:rastrigin}
\end{figure}
%
\begin{table*}
\caption{Summary of fit to GEV and Weibull distribution of
  the number of PUTs per worker for the F15 (Rastrigin) function. \label{tab:puts:ww:f15}}
\begin{center}
\begin{tabular}{cccc}
\hline
Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
\hline
GEV & 35.759  &  343.336   & 9.877 \\
Weibull & ND & 29.002625061 $\pm$ 2.153034198  & 0.402393903 $\pm$ 0.007586468 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{gev-fit-ww-rastrigin-workers.png}
\includegraphics[width=0.49\linewidth]{weibull-fit-ww-rastrigin-workers.png}
\caption{Data and fit to GEV (left) and Weibull (right) of the number
  of PUTs per worker for the F15 (Rastrigin) function.}  
% Plotted with ../data/ips-puts-rastrigin.R
\label{fig:fit:rastrigin}
\end{figure}
%

>>>>

\numberofauthors{5} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
% \author{Juan-J.~Merelo, Dept. of Computer Architecture and Technology
%   and CITIC, University of Granada, Granada, Spain \\
% Mario Garc\'ia-Valdez, Dept. of Graduate Studies, Instituto
% Tecnológico de Tijuana, Tijuana, M\'exico\\
% Pedro A. Castillo, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
% Pablo Garc\'ia-S\'anchez, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\
% Paloma de las Cuevas, Dept. of Computer Architecture and Technology,
% University of Granada, Granada, Spain\\

@@

We were interested in this kind of systems because, during the big
global crisis, our group faced the problem of diminishing funds for
buying new hardware. This was aggravated by the increasing maintenance costs and
extended downtime resulting from the continuous failures of existing
clusters.
Considering this, we leveraged our experience in the design of web
applications since their inception, including JavaScript, and other
volunteer and
unconventional distributed evolutionary computing systems to 
design and release a new free framework that would allow anyone to
create a volunteer distributed evolutionary computation (EC) experiment using cloud resources as
servers, and browsers as clients.
% (Paloma) It seems that you've started the introduction twice. The
% first two paragraphs "introduce the introduction", and so does the
% third. It is specially noticeable when you find "evolutionary
% computation (EC)" twice.
% Yep, I want to say it all, but don't know how to put it together.

@@

%Mario: Rewrote: 
%One of the aspects that must be taken into account in volunteer based
%frameworks is its {\em social} aspect in the design, with 
In general, several volunteers will be needed to
overcome the overhead involved in using a client-server architecture
and this number of volunteers is obviously not guaranteed. Therefore, the {\em social} aspect must be taken into account when designing 
volunteer based frameworks, with 
issues related to security, trust, and privacy among others. 
The computing system becomes then a {\em techno-social system}
\cite{vespignani2009predicting}, in the sense that its social aspects
cannot be separated from its technical specifications, with one
feeding into the other. One of these aspects is safety for the user
running it, as well as for the scientist: the user must be reasonably
sure that the 
experiment is not misusing their resources and the scientist must have a
certain safety in the sense that their server is not going to be
attacked via cheating or denial of service attacks.

We have decided to approach security in a pragmatic way. Including
many security features % (Paloma) I'm kind of lost here... Sentences
                       % seem to be unconnected and unrelated...  
% Separated paragraph and substituting wildcard words - JJ
would decrease performance; on the other hand, we publish all data
resulting from the experiment once private information, such as IPs,
has been processed to an untraceable hash. The experiments themselves
are a proof that there are enough users trusting the experiment to
give it their CPU time and that there has been no attempt to cheat by
sending specially crafted chromosomes or to attack the server to make
it stop working because there was no incentive in doing so.
% (Paloma) Is this really true? Can we really state that there were no
% attacks because there was no incentive? Attackers don't always need
% an incentive to perform an attack, or the incentive could not be
% perceived by us. And why do the experiments prove this? Are there
% specific measures? 
% There were no attacks, that's a fact. There could be in the future,
% but not for the time being - JJ
The
system, besides, is running in a dedicated Platform as a Service (PaaS)
container or Virtual Private Server so accessing the server would only
show the few data obtained from the experiment, adding security against attackers. In addition, making it
free software ensured the meager bounty obtained by attackers were
known by anyone participating in the experiment, via 
links in the web page that hosted it. % (Paloma) What was known? The
                                % fact that it was running in a PaaS?
                                % Why was that known by links?
%That there wasn't anything to hunt. Cleared, I expect.

In this paper, we will carry out three experiments, after establishing
the baseline of the performance that can be expected in a JavaScript
based system and its comparison with other languages in
\cite{2016arXiv160101607Manom}. First, several experiments are
performed using a simple Ajax client that does not 
restart when a solution is found. Second, we change the client a bit
so that it uses web workers, that is, it runs in the background
process and then it restarts when the solution is found. Third, we use the same
% Pablo: it restarts when the solution is found (it has not been
% mentioned yet)
% Just do it! - JJ
client, but with a harder problem that cannot be easily solved by a single
volunteer and thus needs the contribution of many volunteers to
proceed. 
%Mario: refrased: 
%With these three experiments we want mainly to try and model
With these three experiments our main objective is to try and model
the behavior of users and learn how different problems and
implementation choices have an influence on it. That is why in this
paper we will collect information on the number of users participating
in each experiment and also the time they spent doing so. In this
paper we will show how this participation follows certain patterns,
and that they depend on the implementation itself as well as the
problem and its difficulty; however, certain implementations allow to
extract more CPU cycles from volunteers.

@@

%
% \begin{figure}[!htb]
% \centering
% \includegraphics{rastrigin-fitness.png}
% \caption{Fitness of chromosomes sent to the server pool along time in
%   the $x$ axis.} 
% % Plotted with ../data/plot-rastrigin-fitness.R
% \label{fig:puts:rastrigin}
% \end{figure}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{rastrigin-IPs.png}
\includegraphics[width=0.49\linewidth]{rastrigin-IPs-hour.png}
\caption{Number of unique IPs participating in the Rastrigin
  experiment per minute (left) and per hour (right)} 
% Plotted with ../data/plot-rastrigin-IPs.R
\label{fig:ips:rastrigin}
\end{figure}
%
\begin{table}
\caption{Summary of fit to GEV and Weibull distribution of
  the number of PUTs per worker for the F15 (Rastrigin) function. \label{tab:puts:ww:f15}}
\begin{center}
\begin{tabular}{cccc}
\hline
Distribution & Location $\mu$ & Scale $\sigma$ & Shape $\xi$ \\
\hline
GEV & 35.759  &  343.336   & 9.877 \\
Weibull & ND & 29.00 $\pm$ 2.15  & 0.40 $\pm$ 0.01 \\
\hline
\end{tabular}
\end{center}
\end{table}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{gev-fit-ww-rastrigin-workers.png}
\includegraphics[width=0.49\linewidth]{weibull-fit-ww-rastrigin-workers.png}
\caption{Data and fit to GEV (left) and Weibull (right) of the number
  of PUTs per worker for the F15 (Rastrigin) function.}  
% Plotted with ../data/ips-puts-rastrigin.R
\label{fig:fit:rastrigin}
\end{figure}
%
